Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-10-08 manuscript/manuscript.tex
edits

@@

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the impact of this selection
 genome-wide is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change in neutral allele frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, as neutral alleles can remain associated with selected alleles
 over time. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
 one artificial selection experiment in mice, to quantify the effects of linked
 selection over short timescales using covariance among time-points and across
 replicates. We estimate that over \vb{17\% to 37\%} of allele
 frequency change in these experiments is driven by selection in some
 experiments. Against this background of positive genome-wide temporal
 covariances we also identify signals of negative temporal covariance
 corresponding to reversals in the direction of selection for a reasonable
 proportion of loci over the time course of a selection experiment.  Overall,
 we find that in the three studies we analyzed, linked selection has a large
 impact on short-term allele frequency dynamics that is readily distinguishable
 from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work on this question has mostly focused on teasing apart the
impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of a range of sexual species
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,ANDOLFATO,MACPHERSON,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
time.

@@

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
  \end{tabular}
  \caption{A summary of the main selection studies we analyzed.}
\end{table}

@@

\vb{One concern is that these covariances reflect the localized impact of a few
  large-effect loci rather than selection on a polygenic trait. Since our
  covariances are essentially averages over loci, the covariance estimate can
  be strongly affected by a single outlier region. To test whether large
  outlier regions drive the genome-wide signal we see in the
  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
  along the genome (we refer to these as windowed covariances
  throughout) and take the median windowed covariance, and
  trimmed-mean windowed covariance, as a measure of the genome-wide
  covariance robust to large-effect loci. These robust estimates (Supplementary Table \ref{supp:table-trimmed-mean})
  confirm the patterns we see using the mean covariance, confirming that genomic temporal covariances are
 non-zero due to the impact of selection on many genomic windows.

%  This median covariance is also limiting result of a trimmed
 % mean that excludes the upper and lower $\alpha$ tails to calculate the
 % trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
 % trimmed covariance converges to the median windowed covariance (by the
 % definition of the median; see Supplementary Figure
 % \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
 % non-zero due to the impact of selection on many genomic windows.  }

%We
%  take the median windowed covariance as a measure of the genome-wide
 % covariance robust to large-effect loci; we find that the median windowed
 % covariance $\cov(\Delta p_0, \Delta p_{10})$ (pooling all replicates) is also
 % significantly different from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017], see
 % Supplementary Table \ref{supp:table-trimmed-mean} for more timepoints and
 % trimmed means).

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure not easily
interpretable. \vb{We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change:
the ratio of total covariance in allele frequency change to the total variance
in allele frequency change}. We denote the change in allele frequency as
$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
$t$. Since the total variation in allele frequency change can be partitioned
into variance and covariance components, $\var(p_t - p_0) = \sum_{i}
\var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j)$, and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^t \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the effect of selection on allele frequency change between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. As with the temporal covariances, the
study design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even
more conservative, since the temporal covariances within each ten-generation
block between sequenced timepoints are not directly observable, and are not
included in the numerator of $G(t)$. Still, we find a remarkably strong signal.
Greater than $20\%$ of total, genome-wide allele frequency change over 60
generations is the result of selection.

@@

% TODO \vince{GC: you had ??? here, not sure I see the issue.}

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$). Similarly, we can
  calculate the proportion of the total variance in allele frequency change
  from convergent selection pressure analogous to $G$ where the numerator is is
  the convergent covariance and the denominator is the total variance (see
  XXX). We find that 37\% of the total variance is due to shared allele
frequency changes caused by selection (95\% CI [29\%, 41\%]; these are similar
to the convergence correlation \gc{as the variances are relatively
  constant across replicates, see {\bf XXX}).}

@@

We would like to thank the authors of the original studies we've analyzed,
including Neda Barghi, Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes,
Frank Chan, Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  We
would also like to thank Doc Edge for helpful statistical advice, and Matt
Osmond for helpful discussions. This research was supported by an NSF Graduate
Research Fellowship grant awarded to VB (1650042), and NIH (R01-GM108779) and
NSF (1353380) awarded to GC.

@@

\section{Appendix}

\section{Estimator Bias Correction}
\subsection{Correcting variance bias with a single depth sampling process}
\label{supp:depth-var-corr}

@@

\subsection{Correcting variance bias with individual and depth sampling processes}
\label{supp:ind-depth-var-corr}

@@

\subsection{Covariance Correction}
\label{supp:cov-corr}

@@

\subsection{Temporal-Replicate Covariance Matrix Correction}
\label{supp:matrix-correction}

@@

\subsection{\textcite{Barghi2019-qy} Temporal Covariances}
\label{supp:barghi-covs}

@@

\subsection{Block Bootstrap Procedure}
\label{supp:block-bootstrap}

@@

\subsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
\label{supp:replicate-g}

@@

In artificial selection studies with a control line, such as the
\textcite{Castro2019-uk} study, this allows us to estimate the contribution of
the effects of shared and unique indirect selection.

\begin{align}
  \var(\Delta p_{0,\mathrm{Ctrl}}) = \var()
\end{align}

\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
\label{supp:empirical-null}

@@

Since variants have unknown dependency structure due to LD, we sign-permute at
the window level (analogous to our block bootstrap approach). Our covariance
statistic, $Q_{t,s} = \nicefrac{1}{L} \sum_l \Delta p_{t,l} \Delta p_{t,l}$
(for simplicity, we drop the replicate index for clarity, do not center
observations since their expected change is zero, and assume $|t-s| > 1$ to
avoid the need for a bias correction) has a variance that depends on the LD:

\begin{align}
  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \var(\Delta p_{t,l} \Delta p_{s,l}) + \frac{1}{L^2} \sum_{i \ne j} \cov(\Delta p_{t,i}\Delta p_{s,i}, \Delta p_{t,j}\Delta p_{s,j}) 
\end{align}

Under neutrality alone, we can simplify this, using the fact that allele
frequency changes are independent through time,

\begin{align}
  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \E(\Delta p_{t,l}^2) \E(\Delta p_{s,l}^2) + \frac{1}{L^2} \sum_{i \ne j} \E(\Delta p_{t,i} \Delta p_{s,i} \Delta p_{t,j} \Delta p_{t,j}) \\
  &= \frac{1}{L^2} \sum_l \frac{p_{t,l}(1-p_{t,l}) p_{s,l}(1-p_{s,l})}{4N^2} + \\ \nonumber 
  & \;\;\;\;\;\; \frac{1}{L^2} \sum_{i \ne j} \frac{\left(p_{s,i} p_{s,j}-4 D^2 \left(1-2 N r^2\right) \right) \left( p_{t,i} p_{t,j}-4 D^2 \left(1-2 N r^2\right) \right)}{4 N^2}
\end{align}
%

where this last line can be found by assuming haplotype frequencies are sampled
according to a Multinomial process, and recombination occurs after sampling
(see Barton and Otto, 2005). This indicates that the variance of the covariance
statistic, averaged over loci, depends on the linkage disequilibrium term $D^2$
even without the presence of linked selection. However, under neutrality alone,
the covariance statistic is unaffected by LD with sites outside averaged over.
Still, we sign-permute loci at the chromosome level and find that the tail
probabilities do still indicate a reverse in the direction of selection. }

\clearpage
\section{Supplementary Figures}

\subsection{Bias Correction for \textcite{Barghi2019-qy}}

@@

\subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
\gc{ Here we report median and trimmed mean of the windowed covariances. We note that
  the median covariance is also limiting result of a trimmed
  mean that symmetrically excludes the upper and lower $\alpha$ tails to calculate the
 trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
  trimmed covariance converges to the median windowed covariance (by the
 definition of the median; see Supplementary Figure
  \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
  non-zero due to the impact of selection on many genomic windows.  }
 \begin{table}
   \centering
   \footnotesize
   \centering
   \begin{tabular}{l c c c c c c}
     s & t & 5\% CI & median & 95\% CI & trimmed mean \\ \hline

@@

   \end{tabular}
  \label{supp:table-trimmed-mean}
   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
     mean windowed covariance which excludes the lower and upper 5\% windows
     with the highest covariance.}

@@

  \caption{The genome-wide covariance ($\cov(\Delta p_0, \delta p_10)$ pooling
    all replicates) averaged (red line) and the median windowed covariance
    (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed average window
    covariance, excluding the $\alpha$ lower and upper tails, converges to the
     median windowed covariance. This indicates that genome-wide covariance are
     not being overly dominated by a large-effect loci in few windows.}

@@

\clearpage
\subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}

@@

\clearpage
\subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}

@@

\subsection{\textcite{Bergland2014-ij} Re-Analysis}
\label{supp:bergland-reanalysis}

>>>>

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

@@

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the impact of this selection
 genome-wide is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change in neutral allele frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, as neutral alleles can remain associated with selected alleles
 over time. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
 one artificial selection experiment in mice, to quantify the effects of linked
 selection over short timescales using covariance among time-points and across
 replicates. We estimate that at least \vb{17\% to 37\%} of allele frequency
 change in these experiments is driven by selection in these experiments.
 Against this background of positive genome-wide temporal covariances we also
 identify signals of negative temporal covariance corresponding to reversals in
 the direction of selection for a reasonable proportion of loci over the time
 course of a selection experiment. Overall, we find that in the three studies
 we analyzed, linked selection has a large impact on short-term allele
 frequency dynamics that is readily distinguishable from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work on this question has mostly focused on teasing apart the
impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of a range of sexual species
\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
time.

@@

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
  \end{tabular}
  \caption{A summary of the main selection studies we analyzed.}
\end{table}

@@

\vb{One concern is that these covariances reflect the localized impact of a few
  large-effect loci rather than selection on a polygenic trait. Since our
  covariances are essentially averages over loci, the covariance estimate can
  be strongly affected by a single outlier region. To test whether large
  outlier regions drive the genome-wide signal we see in the
  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
  along the genome (we refer to these as windowed covariances throughout) and
  take the median windowed covariance, and trimmed-mean windowed covariance, as
  a measure of the genome-wide covariance robust to large-effect loci. These
  robust estimates (Supplementary Table \ref{supp:table-trimmed-mean}) confirm
the patterns we see using the mean covariance, confirming that genomic temporal
covariances are non-zero due to the impact of selection on many genomic
windows. }

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure not easily
interpretable. \vb{We can calculate a more intuitive measure from the temporal
  covariances to quantify the impact of selection on allele frequency change:
the ratio of total covariance in allele frequency change to the total variance
in allele frequency change}. We denote the change in allele frequency as
$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
$t$. Since the total variation in allele frequency change can be partitioned
into variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i \ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)$ (we bias
correct these for sequencing depth), and the covariances are zero when drift
acts alone, this is a lower bound on how much of the variance in allele
frequency change is caused by linked selection \parencite{Buffalo2019-io}. We
call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the effect of selection on allele frequency change between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. Since \textcite{Barghi2019-qy}
experiment is sequenced every ten generations, in the numerator for the
covariance we use the allele frequency changes between adjacent timepoints,
which are ten generations apart. Consequently, this leads our measure $G(t)$ to
be strongly conservative, since the temporal covariances within each
ten-generation block are not not directly observable, and are not included in
the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
than $20\%$ of total, genome-wide allele frequency change over 60 generations
is the result of selection.

@@

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$). Similarly, we can
calculate the proportion of the total variance in allele frequency change
from convergent selection pressure analogous to $G$ where the numerator is is
the convergent covariance and the denominator is the total variance (see
Supplementary Material \ref{supp:replicate-g}). We find that 37\% of the
total variance is due to shared allele frequency changes caused by selection
(95\% CI [29\%, 41\%]; these are similar to the convergence correlation, see
Supplementary Material \ref{supp:replicate-g}).

@@

We would like to thank the authors of the original studies we've analyzed,
including Neda Barghi, Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes,
Frank Chan, Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  We
would also like to thank Doc Edge for helpful statistical advice, and Matt
Osmond, Erin Calfee, and Sivan Yair for helpful discussions. This research was
supported by an NSF Graduate Research Fellowship grant awarded to VB (1650042),
and NIH (R01-GM108779) and NSF (1353380) awarded to GC.

@@

\setcounter{section}{1}
\section*{Supplementary Material}
\beginsupplement

\subsection{Estimator Bias Correction}
\subsubsection{Correcting variance bias with a single depth sampling process}
\label{supp:depth-var-corr}

@@

\subsubsection{Correcting variance bias with individual and depth sampling processes}
\label{supp:ind-depth-var-corr}

@@

\subsubsection{Covariance Correction}
\label{supp:cov-corr}

@@

\subsubsection{Temporal-Replicate Covariance Matrix Correction}
\label{supp:matrix-correction}

@@

\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances}
\label{supp:barghi-covs}

@@

\subsubsection{Block Bootstrap Procedure}
\label{supp:block-bootstrap}

@@

\subsubsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
\label{supp:replicate-g}

@@

In artificial selection studies with a control (non-selected) line, such as the
\textcite{Castro2019-uk} study, this allows us to estimate the contribution of
the effects of shared and unique indirect selection. In the case of this study,
we can estimate the drift term using the fact that,

\begin{align}
  \Delta p_{t,\mathrm{LS1}} &= \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_\mathrm{LS}} p_t \\ 
  \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}} + \Delta_{_{U}} p_{t,\mathrm{Ctrl}}
\end{align}

\begin{align}
  (\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) = 
  \overline{\var(\Delta_{_D} p_{t,\mathrm{LS}})} + \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
\end{align}

where the bars indicate values averaged

We note that there very well may be selection acting in the control line...

\subsubsection{The Empirical Neutral Null Windowed Covariance Distribution}
\label{supp:empirical-null}

@@

Still, we wanted to ensure that LD between sign-permuted blocks, which will
affect the variance of the empirical null distribution, does not impact our
primary finding that the distribution of temporal covariances becomes
increasingly negative in the \textcite{Barghi2019-qy} dataset through time. To
address this, we also sign-permuted at the whole chromosome level finding we
recapitulate the same pattern (Supplementary Figure
\ref{suppfig:barghi-tailprobs-seqid}).

}

%Since variants have unknown dependency structure due to LD, we sign-permute at
%the window level (analogous to our block bootstrap approach). Our covariance
%statistic, $Q_{t,s} = \nicefrac{1}{L} \sum_l \Delta p_{t,l} \Delta p_{t,l}$
%(for simplicity, we drop the replicate index for clarity, do not center
%observations since their expected change is zero, and assume $|t-s| > 1$ to
%avoid the need for a bias correction) has a variance that depends on the LD:

%\begin{align}
%  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \var(\Delta p_{t,l} \Delta p_{s,l}) + \frac{1}{L^2} \sum_{i \ne j} \cov(\Delta p_{t,i}\Delta p_{s,i}, \Delta p_{t,j}\Delta p_{s,j}) 
%\end{align}

%Under neutrality alone, we can simplify this, using the fact that allele
%frequency changes are independent through time,

%\begin{align}
%  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \E(\Delta p_{t,l}^2) \E(\Delta p_{s,l}^2) + \frac{1}{L^2} \sum_{i \ne j} \E(\Delta p_{t,i} \Delta p_{s,i} \Delta p_{t,j} \Delta p_{t,j}) \\
%  &= \frac{1}{L^2} \sum_l \frac{p_{t,l}(1-p_{t,l}) p_{s,l}(1-p_{s,l})}{4N^2} + \\ \nonumber 
%  & \;\;\;\;\;\; \frac{1}{L^2} \sum_{i \ne j} \frac{\left(p_{s,i} p_{s,j}-4 D^2 \left(1-2 N r^2\right) \right) \left( p_{t,i} p_{t,j}-4 D^2 \left(1-2 N r^2\right) \right)}{4 N^2}
%\end{align}
%%

%where this last line can be found by assuming haplotype frequencies are sampled
%according to a Multinomial process, and recombination occurs after sampling
%(see Barton and Otto, 2005). This indicates that the variance of the covariance
%statistic, averaged over loci, depends on the linkage disequilibrium term $D^2$
%even without the presence of linked selection. However, under neutrality alone,
%the covariance statistic is unaffected by LD with sites outside averaged over.
%Still, we sign-permute loci at the chromosome level and find that the tail
%probabilities do still indicate a reverse in the direction of selection. }

\clearpage

\setcounter{section}{1}
\section*{Supplementary Figures}

\subsubsection{Bias Correction for \textcite{Barghi2019-qy}}

@@

\subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}

Here we report median and trimmed mean of the windowed covariances. We note
that the median covariance is also limiting result of a trimmed mean that
symmetrically excludes the upper and lower $\alpha$ tails to calculate the
trimmed average windowed covariance. As $\alpha$ increases to 0.5, the trimmed
covariance converges to the median windowed covariance (by the definition of
the median; see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus
our genomic temporal covariances are non-zero due to the impact of selection on
many genomic windows.

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    s & t & 5\% CI & median & 95\% CI & trimmed mean \\ \hline

@@

   \end{tabular}
   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
     mean windowed covariance which excludes the lower and upper 5\% windows
     with the highest covariance.}
  \label{supp:table-trimmed-mean}

@@

  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_10)$ pooling
    all replicates) averaged (red line) and the median windowed covariance
    (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed average window
    covariance, excluding the $\alpha$ lower and upper tails, converges to the
     median windowed covariance. This indicates that genome-wide covariance are
     not being overly dominated by a large-effect loci in few windows.}

@@

\clearpage
\subsubsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}

@@

\clearpage
\subsubsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}

@@

\subsubsection{\textcite{Bergland2014-ij} Re-Analysis}
\label{supp:bergland-reanalysis}
